---
output: github_document
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/UrbanInstitute/nccsdata/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/UrbanInstitute/nccsdata/actions/workflows/R-CMD-check.yaml)
[![test-coverage](https://github.com/UrbanInstitute/nccsdata/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/UrbanInstitute/nccsdata/actions/workflows/test-coverage.yaml)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nccsdata

## Overview

nccsdata provides tools to read, filter and append metadata to publicly available NCCS Core and BMF data sets.

## Installation

You can install the development version of nccsdata from [GitHub](https://github.com/) with:

``` {r, message = FALSE, eval = FALSE}
install.packages("devtools")
devtools::install_github("UrbanInstitute/nccsdata")
library(nccsdata)
```

## Usage

### Data Pulls

The [`nccsdata`](https://urbaninstitute.github.io/nccsdata/) package can be used to download legacy core data from 1989 to 2019 for charities, nonprofits, or private foundations that file their respective required IRS forms such as Form 990, 990EZs, or both.

This data can be filtered based on [NTEE](https://github.com/Nonprofit-Open-Data-Collective/mission-taxonomies/blob/main/NTEE-disaggregated/README.md) codes and geography.

```{r example, message=FALSE}
core_2005_nonprofit_pz <- nccsdata::get_data(dsname = "core",
                                             time = "2005",
                                             scope.orgtype = "NONPROFIT",
                                             scope.formtype = "PZ")


tibble::as_tibble(core_2005_nonprofit_pz)
```

``` {r message = FALSE, warning = FALSE}
core_2005_artnonprofits_newyork <- nccsdata::get_data(dsname = "core",
                                                      time = "2016",
                                                      scope.orgtype = "NONPROFIT",
                                                      scope.formtype = "PZ",
                                                      ntee = "ART",
                                                      geo.state = "NY")
tibble::as_tibble(core_2005_artnonprofits_newyork)
```

 * Full [`get_data()`](https://urbaninstitute.github.io/nccsdata/articles/data_pull.html) vignette

### Summarising Data

After processing the desired data, [`nccsdata`](https://urbaninstitute.github.io/nccsdata/) can also be used to 
generate summary tables.

```{r message = FALSE, warning = FALSE}
nccsdata::preview_sample(data = core_2005_artnonprofits_newyork,
                         group_by = c("NTEECC", "STATE"),
                         var = c("TOTREV"),
                         stats = c("count", "mean", "max"))
```

 * Full [`preview_sample()`](https://urbaninstitute.github.io/nccsdata/articles/summary_stats.html) vignette.
 
### NTEE Codes
 
 [`nccsdata`](https://urbaninstitute.github.io/nccsdata/) also offers several
 supplementary functions for documenting and retrieving [NTEE](https://github.com/Nonprofit-Open-Data-Collective/mission-taxonomies/blob/main/NTEE-disaggregated/README.md) codes.
 
  * Full [`ntee_preview()` and `parse_ntee()`](https://urbaninstitute.github.io/nccsdata/articles/ntee.html) vignette.


## Getting Help

Raise an issue on the [issues](https://github.com/UrbanInstitute/nccsdata/issues) page or contact Thiyaghessan at `tpoongundranar@urban.org`.
