---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# nccsdata <img src="hex/nccsdata.svg" align="right" height="139" alt="nccsdata hex logo" />

<!-- badges: start -->
[![R-CMD-check](https://github.com/UrbanInstitute/nccsdata/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/UrbanInstitute/nccsdata/actions/workflows/R-CMD-check.yaml)
[![test-coverage](https://github.com/UrbanInstitute/nccsdata/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/UrbanInstitute/nccsdata/actions/workflows/test-coverage.yaml)
<!-- badges: end -->

nccsdata provides tools to download, filter, and analyze nonprofit organization data from the [National Center for Charitable Statistics](https://nccs.urban.org/) (NCCS). It reads IRS Business Master File (BMF) data stored as parquet files in a public S3 bucket, with support for predicate-pushdown filtering by state, county, NTEE subsector, and exempt organization type.

> **Note:** This is version 2.0.0, a ground-up rewrite of the package. The v1 API (`get_data()`, `preview_sample()`, `parse_ntee()`) has been replaced. See the [migration section](#migrating-from-v1) below.

## Installation

Install the development version from GitHub:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("UrbanInstitute/nccsdata")
```

## Usage

### Reading BMF data

`nccs_read()` downloads BMF data from S3 with optional filters. Filtering happens at the Arrow level via predicate pushdown, so only matching rows are read into memory.

```{r, eval = FALSE}
library(nccsdata)

# All Pennsylvania nonprofits (default columns)
pa <- nccs_read(state = "PA")

# Arts nonprofits in New York
ny_arts <- nccs_read(state = "NY", ntee_subsector = "ART")

# Select specific columns
pa_slim <- nccs_read(
  state = "PA",
  columns = c("ein", "org_name_display", "geo_county", "income_amount")
)

# Lazy query for custom dplyr pipelines
query <- nccs_read(state = "PA", collect = FALSE)
result <- query |>
  dplyr::filter(geo_county == "Lackawanna County") |>
  dplyr::collect()
```

### Summarizing data

`nccs_summary()` produces grouped count summaries from a collected data frame.

```{r, eval = FALSE}
pa <- nccs_read(state = "PA")

# Total count
nccs_summary(pa)

# Count by county
nccs_summary(pa, group_by = "geo_county")

# Count by county and subsector, export to CSV
nccs_summary(pa, group_by = c("geo_county", "nteev2_subsector"),
             output_csv = "pa_counts.csv")
```

### Discovering valid filter values

`nccs_catalog()` lists valid values for `nccs_read()` filters without any network calls.

```{r, eval = FALSE}
nccs_catalog("state")
nccs_catalog("ntee_subsector")
nccs_catalog("exempt_org_type")

# Pass `labels = TRUE` for a code + description tibble, sourced from the
# bundled BMF lookup tables.
nccs_catalog("ntee_subsector", labels = TRUE)
nccs_catalog("foundation_code", labels = TRUE)
```

### Cleaning external data

The BMF returned by `nccs_read()` is already normalized upstream, but two helpers are exposed for users joining external CSVs or API responses against it:

```{r, eval = FALSE}
# Coerce EINs in any format to canonical XX-XXXXXXX
nccs_normalize_ein(c("123456789", "12-3456789", 1234567))
#> [1] "12-3456789" "12-3456789" "00-1234567"

# Coerce IRS binary-indicator columns to logical
nccs_as_indicator(c("Y", "N", "1", "2"))
#> [1]  TRUE FALSE  TRUE FALSE

# e-file indicator accepts E/P (2015, 2018+) and Y/N (2016-2017)
nccs_as_indicator(c("E", "P", "Y", "N"), scheme = "efile")
```

### Browsing the data dictionary

`nccs_dictionary()` returns a tibble describing all BMF columns, with optional pattern filtering.

```{r, eval = FALSE}
# All columns
nccs_dictionary()

# Find geocoding-related columns
nccs_dictionary("geo")

# Find NTEE-related columns
nccs_dictionary("ntee")
```

## Scope and design

`nccsdata` is intentionally a lean reader. A few principles that shape what is — and is not — in the package:

- **No re-cleaning of upstream data.** The BMF and CORE Series parquet files are cleaned by the sibling ETL pipelines (`nccs-data-bmf`, `nccs-data-core`). EIN normalization, NTEE decoding, geocoding, and subsection labeling are done before publication. We don't re-implement them here.
- **The two exceptions are helpers for *external* data.** `nccs_normalize_ein()` and `nccs_as_indicator()` exist so you can bring your own CSVs (member rosters, survey extracts, donor lists) into the same shape as the package's output before joining.
- **One opinionated analytic helper: inflation adjustment.** `nccs_deflate()` and the bundled annual `cpi_u` series are included because real-dollar conversion needs a reference table the user otherwise has to fetch themselves, and the conversion itself is mechanical and uncontroversial.
- **No canonical financial ratios.** Operating margin, program-expense ratio, fundraising efficiency, months of operating reserves, and similar measures are deliberately *not* bundled. Their definitions vary by analyst (which numerator, which denominator, which exclusions), and shipping one canonical version would make the package take editorial sides. They're also one-line `mutate()` calls on the columns CORE already provides.
- **Lean dependencies.** Hard imports are `arrow`, `dplyr`, `utils`. Anything heavier (sf, tigris, ggplot2, data.table) belongs in vignettes that show how to combine `nccsdata` with those packages, not as a dependency.

If you want to build analytic functionality on top of this package, the right pattern is a downstream package or notebook that imports `nccsdata` and adds your team's preferred ratio definitions.

## Migrating from v1

| v1 function | v2 replacement |
|---|---|
| `get_data()` | `nccs_read()` |
| `preview_sample()` | `nccs_summary()` |
| `ntee_preview()` / `parse_ntee()` | `nccs_catalog("ntee_subsector")` |

Key changes:

- Data source moved from legacy Core/BMF CSVs to geocoded BMF parquet files on S3.
- Filtering now uses Arrow predicate pushdown instead of downloading full files.
- Dependencies reduced from 12 packages to 3 (`arrow`, `dplyr`, `utils`).

## Documentation

Full documentation is available at <https://urbaninstitute.github.io/nccsdata/>.

## Getting help

- Browse the [getting started vignette](https://urbaninstitute.github.io/nccsdata/articles/getting-started.html)
- Open an issue on [GitHub](https://github.com/UrbanInstitute/nccsdata/issues)
- Contact the maintainer at `tpoongundranar@urban.org`
