---
title: "Data Downloads with nccsdata"
date: '`r format(Sys.Date(), "%B %d %Y")`'
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
    vignette: >
      %\VignetteIndexEntry{valr-interval-statistics}
      %\VignetteEngine{knitr::rmarkdown}
      %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
```

## Overview

`nccsdata` includes a core get_data() function that downloads data from NCCS'
publicly accessible S3 buckets and processes them for the user.

In this vignette, we provide several examples of how this function can be used
to retrieve NCCS data.

```{r load-data, message = FALSE, warning = FALSE}

library(nccsdata)
```

## Downloading Core Data

First, we can define the type of data and the years for which that data is
available using the `dsname` and `time` arguments.

```{r}
core <- get_data(dsname = "core",
                 time = "2015")
```

The function downloads NCCS core data for the year 2015.
