---
title: "Summarising nccsdata"
date: '`r format(Sys.Date(), "%B %d %Y")`'
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
    vignette: >
      %\VignetteIndexEntry{Summarising nccsdata}
      %\VignetteEngine{knitr::rmarkdown}
      %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
```

## Overview

```{r load-data, message = FALSE, warning = FALSE}
library(nccsdata)
```

After preparing data with `get_data()`, `preview_sample()` can be used to
compute summary statistics and print the results in a tabular format for 
different subgroups in the prepared data.

In this vignette, we will provide examples of `preview_sample()`.

## Pulling Data and Computing Summary Statistics

```{r warning = FALSE}
core <- get_data(dsname = "core",
                 time = "2015")
preview_sample(data = core,
              group_by = c("NTEECC", "STATE"),
              var = c("TOTREV"),
              stats = c("count", "mean", "max"))
```

`preview_sample()` groups the data set by user-defined `group_by` columns, and computes summary statistics for the user-defined `var` column. The available
summary statistics are:

 * `min`: minimum value
 * `median`: median value
 * `max`: maximum value
 * `mean`: mean value
 * `count`: count of rows belonging to group
 


